Optimize the Cannon 2

### 2016-10-16 paper/ms.tex
Add fitting spectrum

% ****** Start of file apssamp.tex ******
%
%   This file is part of the APS files in the REVTeX 4.1 distribution.
%   Version 4.1r of REVTeX, August 2010
%
%   Copyright (c) 2009, 2010 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.1
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
\documentclass[%
 reprint,
%superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%showpacs,preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps,
%pra,
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-1}

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

%\usepackage[showframe,%Uncomment any one of the following lines to test 
%%scale=0.7, marginratio={1:1, 2:3}, ignoreall,% default settings
%%text={7in,10in},centering,
%%margin=1.5in,
%%total={6.5in,8.75in}, top=1.2in, left=0.9in, includefoot,
%%height=10in,a5paper,hmargin={3cm,0.8in},
%]{geometry}

\begin{document}

\preprint{APS/123-QED}

\title{Optimized Synthesized Spectrum}% Force line breaks with \\

\author{Jason Cao}

\affiliation{%
 Physics Department, New York University
}%

\date{\today}% It is always \today, today,
             %  but any date may be explicitly specified

\begin{abstract}
The Cannon 2 is a powerful tool, which can infer labels from the spectrum after it's trained. Also, it can generate the spectrum. Here we start at the point that the model is trained. First, infer the spectrum by using the Cannon 2. Then optimize the method and generate a better spectrum. Finally, compare the synthesized spectrum with the real spectrum and check whether the optimization works.
\begin{description}

\item[Structure]
There are four sections. The first part is the introduction part, which describes the background of the method. Then in the assumptions and method part, we show the detail of our method. The next section is the experiment, where the experiment results are showed. Finally, we discuss our results?
\end{description}
\end{abstract}

\pacs{}% PACS, the Physics and Astronomy
                             % Classification Scheme.
%\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle

%\tableofcontents

@@

 \section{\label{sec:level1}Assumption and method}

Suppose we have a trained model. The training set contains 548 stellar and each stellar spectrum has 8575 pixels.

@@

 Where \(y_{jn}\) is the spectrum data for star n at wavelength pixel j. 
 And \(v(l_n)\) is the vectorizing function. 
The input \(l_n\) is the label list of length K for star n and the  output 
\(v(l_n)\) is a vector of length D (D is bigger than K).
\( \theta_{j}\) is a vector of length D of parameters which controlling the model at wavelength pixel j.
\( e_{jn}\) is a noise draw or residual at pixel j for star n.

The inferred label for stellar n \(l_n^{inferred}\) and \( \theta_{j}\) are available after the training step.
Then infer spectrum by using:
 \[y_{jn}^{inferred} = v(l_n^{inferred}) \cdot \theta_{j}  \]

 This is how to obtain the synthesized spectrum by using the Cannon 2.
 The code is as followed:

 \bigskip
inferred\_labels = model.fit\_labelled\_set()

 \bigskip
v = model.vectorizer.get\_label\_vector(inferred\_labels)
\bigskip

inferred\_flux = np.dot(v,model.theta.T)
 \bigskip

@@

From the inferred spectrum \(y_{jn}^{inferred}\), we optimize it and generate a better synthesized spectrum, which we call \(y_{jn}^{optimized}\).
\bigskip

The method we use is to represent the spectrum at pixel j by using the inferred spectrum at pixel j-1,pixel j and pixel j+1.(If the pixel is out of range, set the flux as 1). 
\bigskip

Because linear regression is sensitive to bad data points, extra parameters need to be added to describe the bad data points. Here we introduce \(\big\{q_n\big\}_{n=1}^{N}\), \(P\), \(Y\) and \(V\).

We need to find a correct objective function! 
\[y_{j,n}^{optimized}=a\cdot y_{j-1,n}^{inferred}+b\cdot y_{j,n}^{inferred}+c\cdot y_{j+1,n}^{inferred}(2)\]
\bigskip

Then fit \(y_{j,n}^{optimized} \) with \(y_{jn}\), where \(y_{jn}\) is the spectrum data for star n at wavelength pixel j. The optimization method we use is Levenberg - Marquardt algorithm.
a+b+c should be equal to 1.
\bigskip

The code is as followed:
\bigskip

popt, pcov= optimization.curve\_fit(func, (x\_data, y\_data, z\_data),nor\_r, x0)
\bigskip

parameters = popt
\bigskip

x\_data, y\_data and z\_data are \(y_{j-1,n}^{inferred},y_{j,n}^{inferred} \)

and \( y_{j+1,n}^{inferred}\)
\bigskip

parameters are a, b and c

The result is \(a+b+c = 0.982\). (a,b,c) = ()

Calculate the optimized spectrum \( y_{j,n}^{optimized}\) by using (2)
\bigskip

If the optimization works, we will optimized \(\theta_{j}\) besides a,b and c, which will come soon.

@@

 Choose stellar aspcapStar-v304-2M16421116+361822
 5. Then plot the spectrum

 The blue line "Observed" is the observed spectrum from the data  \(y_{jn}\).
 The green line "inferred" is \( y_{j,n}^{inferred}\) , which is generated by the Cannon 2.
 The red line "optimized" is \( y_{j,n}^{optimized}\), which is generated by (2)

%\includegraphics[width=70mm]{figure_1.png}

 Then compare the chi-square for aspcapStar-v304-2M16421116+3618225 of the two methods:
 chi-square for \( y_{j,n}^{inferred}\) and \( y_{j,n}^{optimized}\) are 302.41 and 304.21
 They are similar to each other.

@@

\end{document}
%
% ****** End of file apssamp.tex ******

>>>>

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.

\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}
\usepackage{amsmath, amsthm, amssymb, amsfonts}

\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{authblk}

%SetFonts

%SetFonts

\title{Optimize the Cannon 2}
\author[1]{Jason Cao\thanks{jc6933@nyu.edu}}
\author[2]{David W. hogg\thanks{david.hogg@nyu.edu}}

\affil[1]{Department of Physics,  New York University}
\affil[2]{NYU Physics - Center for Cosmology and Particle Physics
NYU Center for Data Science
Max-Planck-Institut fuer Astronomie }

\renewcommand\Authands{ and }

\date{\today}						% Activate to display a given date or no date

\begin{document}
\maketitle

@@

 \section{\label{sec:level1}Assumption and method}
 \subsection{\label{sec:level2}Optimize the spectrum}

Suppose we have a trained model. The training set contains 548 stellar and each stellar spectrum has 8575 pixels. Every pixel is related to a specific wave length. It belongs to APOGEE data release 10 and has a good quality. These stellar are the very same objects as used by the APOGEE survey for the calibration of DR10. Each of them has at least nine stellar labels. But in the training step, we only use three of them $T_{eff}$, $log$ g  and $\left[Fe/H\right]$. They span the range of $3500K <T_{eff}< 5300K$, $0 <log$ $g< 5$ and $-2.5 <\left[Fe/H\right]< 0.45$.
]

@@

 Where \(y_{jn}\) is the spectrum data for star n at wavelength pixel j. 
 \(v(l_n)\) is the vectorizing function. 
The input \(l_n\) is the label list of length K for star n and the  output 
\(v(l_n)\) is a vector of length D (D is bigger than K).
\( \theta_{j}\) is a vector of length D of parameters which controlling the model at wavelength pixel j.
\( e_{jn}\) is a noise draw or residual at pixel j for star n.

\(l_n^{inf}\) and \( \theta_{j}\) are available after the training step.
Infer spectrum by using:
 \[y_{jn}^{inf} = v(l_n^{inf}) \cdot \theta_{j}  (2)\]

 This is how to obtain the synthesized spectrum by using the Cannon 2. We also call it inferred flux(spectrum).

@@

From the inferred spectrum \(y_{jn}^{inf}\), we optimize it and generate a better synthesized spectrum, which we call \(y_{jn}^{opt}\).
\bigskip

The method we use is to represent the spectrum at pixel j by using the inferred spectrum at pixel j-1,pixel j and pixel j+1.(If the pixel is out of range, set the flux as 1). The data we use is the same as the training step.
\bigskip

The new spectrum, which is the optimized spectrum, is represented by:
\[y_{j,n}^{opt}=a\cdot y_{j-1,n}^{inf}+b\cdot y_{j,n}^{inf}+c\cdot y_{j+1,n}^{inf}(3)\]
\bigskip

We use the least chi-squared method. So our objective function is the chi-squared between the optimized spectrum and the spectrum data. 
\bigskip

$F_{obj} (a,b,c)$= \[\sum_{n=1,j=1}^{n=N,j=J} {(y_{jn} - y_{jn}^{opt})^2 \over{\sigma_{jn}^2}} \]  (4)
\bigskip

$\sigma_{jn}^2$ is the data uncertainty for star n at pixel j. The number of stellar and pixel each star are N and J. Here N=548 and J = 8575. Write the objective function in the form of matrix can make it simpler. So we introduce several matrix: Y, A, C and X, which are defined as followed:

% insert matrix

\[
Y
=
\begin{bmatrix}
    Y_1&Y_2 &\dots&Y_{N}\\

\end{bmatrix}
\]

where
\[
 Y_n
=
\begin{bmatrix}
    y_{1,n} \\
    y_{2,n} \\
    \vdots  \\
    y_{J,n}  \\

\end{bmatrix}
\]

% A
\[
A
=
\begin{bmatrix}
    A_1&A_2 &\dots&A_{N}\\

\end{bmatrix}
\]

where
\[
A_n
=
\begin{bmatrix}

    1 & y_{1,n} ^{inf} & y_{2,n} ^{inf}\\
    y_{1,n}^{inf}&y_{2,n}^{inf}& y_{3,n}^{inf}  \\
    \vdots & \vdots & \vdots \\
    y_{J-1,n}^{inf}&y_{J,n}^{inf}& 1  \\

\end{bmatrix}
\]

% sigma-C

\[
C
=
\begin{bmatrix}
    C_1&C_2 &\dots&C_{N}\\

\end{bmatrix}
\]

where
\[
C_n
=
\begin{bmatrix}

    \sigma_{1,n}^2&0&\dots&0\\
    0& \sigma_{2,n}^2&\dots&0 \\
    \vdots & \vdots & \vdots \\
    0&\dots&0&\sigma_{J,n}^2  \\

\end{bmatrix}
\]

\bigskip

\(X = 
 \begin{bmatrix}

    a\\
    b \\
    c \\

\end{bmatrix}
\)

We cut matrix Y, A and C into N smaller matrix so the calculation speed is higher. 
\bigskip

Now the objective function can be described as:
$F_{obj} (a,b,c)$= \[\sum_{n=1}^{N} (Y_n-A_n X)^TC^{-1}(Y_n-A_n X) (5)\]  
\bigskip

And the Jacobian Matrix of the objective function with respect to parameters is

J = $\partial F_{obj}\over \partial X$ \[=\sum_{n=1}^{n=N}2\cdot [(Y_n-A_nX)^TC^{-1}]\cdot (-A_n) (6)\]   
\bigskip

The dimension of the Jacobian Matrix is 1*3. To minimize the objective function, let the Jacobian Matrix be 0 and we have:

\[X = [\sum_{n=1}^{n=N}A_n^TC_n^{-1}A]^{-1} \cdot [\sum_{n=1}^{n=N}A^TC_n^{-1}Y_n](7)\]
\bigskip

By using (7) we can calculate a,b and c. b should be much bigger than both a and c. And a+b+c should be equal to 1.

The result is \(a+b+c = 0.99990\).

(a,b,c) = (-0.00215, 1.00462, -0.00258 ).

Calculate the optimized spectrum \( y_{j,n}^{opt}\) by using (3). The Experiment 1 shows that the optimization works.
\bigskip

%If the optimization works, we will optimized \(\theta_{j}\) besides a, b and c, which will come soon.
%It works!

 \subsection{\label{sec:level2}Optimized $\theta_{j}$ and inferred labels}
 After optimizing the inferred spectrum, we can modify $\theta_{j}$ and get a better result. The method is to represent $\theta_j$ by using $\theta_{j-1}$ , $\theta_j$ and $\theta_{j+1}$. If $\theta_j$ is out of range, set it as zero. So the optimized $\theta$ is:
 \[\theta_{j}^{opt}=a\cdot \theta_{j-1}+b\cdot \theta_{j}+c\cdot \theta_{j+1}(8)\]
 \bigskip

 Here a,b and c are the parameters obtained from spectrum optimization. Then we infer stellar labels by using $\theta_{j}^{opt}$ and generate optimized inferred labels. The data we use is the training set, which includes 548 stellar with 8575 pixel each.
 Finally, compare the chi-squared to check whether the optimization works.

@@

 Choose four stellar, which are not in the training set:

 Star A  $T_{eff} =4750$  log g =3.0 $\left[Fe/h\right] =0.15$

 Star B  $T_{eff} =4849$  log g =2.2 $\left[Fe/h\right] =-1.0$

 Star C  $T_{eff} =3614$  log g =0.4 $\left[Fe/h\right] =-0.68$

 Star D  $T_{eff} =5003$  log g =2.8 $\left[Fe/h\right] =-0.71$

 Then plot the spectrum

 The black line "Observed" is the observed spectrum from the data  \(y_{jn}\).
 The green line "inferred" is \( y_{j,n}^{inf}\) , which is generated by the Cannon 2.
 The red line "optimized" is \( y_{j,n}^{opt}\), which is generated by (3)
\bigskip

%\includegraphics[width=170mm]{figure_1.png}
 \centerline{Fig.1 The optimized spectrum and the data spectrum}
\bigskip

%\includegraphics[width=170mm]{figure_2.png}
 \centerline{Fig.2 The inferred spectrum and the data spectrum}

 Finally compare the chi-square for these four stars of the two methods:
 The optimized spectrum has a 1\% smaller chi-squared so our optimization works.

 \subsection{\label{sec:level2}Experiment 2}
 Optimize $\theta_j$ and generate optimized inferred labels. Then plot the inferred labels and the optimized inferred labels.
 \bigskip

 %\includegraphics[width=170mm]{figure_3.png}
 \centerline{Fig.3 Inferred labels and the reference labels}
 \bigskip

 Finally compare the chi-square.

The chi-squared becomes 1\% bigger after the optimization. We need to do bigger optimization before inferring the labels!

@@

\end{document}
